Oral Manuscript

### 2018-06-20 Manuscript.tex
Deploy MGallow/oral_manuscript to github.com/MGallow/oral_manuscript.git:gh-pages

\documentclass[12pt,]{book}

@@

\title{Oral Manuscript}
\author{Matt Galloway}
\date{2018-06-07}

@@

\begin{align*}
  \mbox{minimize } f(x) + g(z) \\
  \mbox{subject to } Ax + Bz = c
\end{align*}

@@

\hypertarget{references}{%
\section{References}\label{references}}

@@

\hypertarget{compound-symmetric-p-100-n-50}{%
\subsection{Compound Symmetric: P = 100, N =
50}\label{compound-symmetric-p-100-n-50}}

@@

\hypertarget{compound-symmetric-p-10-n-1000}{%
\subsection{Compound Symmetric: P = 10, N =
1000}\label{compound-symmetric-p-10-n-1000}}

@@

\hypertarget{dense-p-100-n-50}{%
\subsection{Dense: P = 100, N = 50}\label{dense-p-100-n-50}}

@@

\hypertarget{dense-p-10-n-50}{%
\subsection{Dense: P = 10, N = 50}\label{dense-p-10-n-50}}

@@

\hypertarget{tridiagonal-p-100-n-50}{%
\subsection{Tridiagonal: P = 100, N = 50}\label{tridiagonal-p-100-n-50}}

@@

>>>>

\documentclass[11pt,]{book}

@@

\title{Oral Manuscript}
\author{Matt Galloway}
\date{2018-06-20}

@@

\[
\begin{aligned}
  \mbox{minimize } f(x) + g(z) \\
  \mbox{subject to } Ax + Bz = c
\end{aligned}
\]

@@

@@

\hypertarget{compound-symmetric-p-100-n-50}{%
\section{Compound Symmetric: P = 100, N =
50}\label{compound-symmetric-p-100-n-50}}

@@

\hypertarget{compound-symmetric-p-10-n-1000}{%
\section{Compound Symmetric: P = 10, N =
1000}\label{compound-symmetric-p-10-n-1000}}

@@

\hypertarget{dense-p-100-n-50}{%
\section{Dense: P = 100, N = 50}\label{dense-p-100-n-50}}

@@

\hypertarget{dense-p-10-n-50}{%
\section{Dense: P = 10, N = 50}\label{dense-p-10-n-50}}

@@

\hypertarget{tridiagonal-p-100-n-50}{%
\section{Tridiagonal: P = 100, N = 50}\label{tridiagonal-p-100-n-50}}

@@

\hypertarget{penalized-regression-estimators}{%
\chapter{Penalized Regression
Estimators}\label{penalized-regression-estimators}}

This document explores a number of regression estimators subject to
various assumptions. We will assume \(n\) samples of

\[ Y_{i} = \mu_{y} + \beta^{T}(X_{i} - \mu_{x}) + \epsilon_{i} \]

where \(\epsilon_{i} \sim N_{r}\left(0, \Sigma_{y | x} \right)\) and
\(X_{i} \sim N_{p}\left(0, \Sigma_{x}\right)\) so that by taking
\(\theta = \left( \mu_{y}, \mu_{x}, vec\left( \beta \right), vec\left( \Sigma_{y | x}^{-1} \right), vec\left( \Sigma_{xx}^{-1} \right) \right)^{T}\)
the joint log-likelihood is of the following form:

\begin{align*}
  l(\theta) &= \frac{nr}{2}log(2\pi) - \frac{n}{2}log\left| \Sigma_{y | x}^{-1} \right| + \frac{1}{2}\sum_{i = 1}^{n}\left( Y_{i} - \mu_{y} - \beta^{T}\left( X_{i} - \mu_{x} \right) \right)^{T}\Sigma_{y | x}^{-1}\left( Y_{i} - \mu_{y} - \beta^{T}\left( X_{i} - \mu_{x} \right) \right) \\
  &+ \frac{np}{2}log(2\pi) - \frac{n}{2}log\left| \Sigma_{xx}^{-1} \right| + \frac{1}{2}\sum_{i = 1}^{n}\left( X_{i} - \mu_{x} \right)^{T}\Sigma_{xx}^{1}\left( X_{i} - \mu_{x} \right) \\
\end{align*}

For convenience, we will later denote \(\mathbb{X}\) as the
\(n \times p\) matrix with rows \(X_{i} - \bar{X}\) where
\(\bar{X} = \sum_{i = 1}^{n}X_{i}/n\) and \(\mathbb{Y}\) as the
\(n \times r\) matrix with rows \(Y_{i} - \bar{Y}\) and \(\bar{Y}\)
defined similarly. Note that \(\bar{X}\) and \(\bar{Y}\) are the maximum
likelihood esitmators of \(\mu_{x}\) and \(\mu_{y}\), respectively.

\hypertarget{formulas}{%
\section{Formulas}\label{formulas}}

Below we outline a few important formulas that will be used throughout
this report:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Forward regression coefficient:
  \(\beta = \Sigma_{xx}^{-1}\Sigma_{xy} = \Sigma_{x | y}^{-1}\alpha^{T}\left( \Sigma_{yy}^{-1} + \alpha\Sigma_{x | y}^{-1}\alpha^{T} \right)^{-1}\)
\end{enumerate}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  Inverse regression coefficient:
  \(\alpha = \Sigma_{yy}^{-1}\Sigma_{xy}^{T}\)
\end{enumerate}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  \(\Sigma_{y | x} = \Sigma_{yy} - \beta^{T}\Sigma_{xx}\beta = \Sigma_{yy} - \Sigma_{xy}^{T}\Sigma_{xx}^{-1}\Sigma_{xy}\)
\end{enumerate}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{3}
\tightlist
\item
  \(\Sigma_{y | x}^{-1} = \Sigma_{yy}^{-1} + \Sigma_{y | x}^{-1}\Sigma_{xy}^{T}\left( \Sigma_{yy} + \Sigma_{xy}\Sigma_{y | x}^{-1}\Sigma_{xy}^{T} \right)^{-1}\Sigma_{xy}\Sigma_{y | x}^{-1} = \Sigma_{yy}^{-1} + \Sigma_{y | x}^{-1}\beta\left( \Sigma_{xx}^{-1} + \beta\Sigma_{y | x}^{-1}\beta^{T} \right)^{-1}\beta\Sigma_{y | x}^{-1}\)
\end{enumerate}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{4}
\tightlist
\item
  \(\Sigma_{x | y} = \Sigma_{xx} - \alpha^{T}\Sigma_{yy}\alpha = \Sigma_{xx} - \Sigma_{xy}\Sigma_{yy}^{-1}\Sigma_{xy}^{T}\)
\end{enumerate}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{5}
\tightlist
\item
  \(\Sigma_{x | y}^{-1} = \Sigma_{xx}^{-1} + \Sigma_{x | y}^{-1}\Sigma_{xy}\left( \Sigma_{yy} + \Sigma_{xy}^{T}\Sigma_{x | y}^{-1}\Sigma_{xy} \right)^{-1}\Sigma_{xy}^{T}\Sigma_{x | y}^{-1} = \Sigma_{xx}^{-1} + \Sigma_{x | y}^{-1}\alpha\left( \Sigma_{yy}^{-1} + \alpha\Sigma_{x | y}^{-1}\alpha^{T} \right)^{-1}\alpha\Sigma_{x | y}^{-1}\)
\end{enumerate}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{6}
\tightlist
\item
  \(\Sigma_{yy}^{-1} = \Sigma_{y | x}^{-1} - \Sigma_{y | x}^{-1}\Sigma_{xy}^{T}\left( \Sigma_{xx} + \Sigma_{xy}\Sigma_{y | x}^{-1}\Sigma_{xy}^{T} \right)^{-1}\Sigma_{xy}\Sigma_{y | x}^{-1} = \Sigma_{y | x}^{-1} - \Sigma_{y | x}^{-1}\beta\left( \Sigma_{xx}^{-1} + \beta\Sigma_{y | x}^{-1}\beta^{T} \right)^{-1}\beta\Sigma_{y | x}^{-1}\)
\end{enumerate}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{7}
\tightlist
\item
  \(\Sigma_{xx}^{-1} = \Sigma_{x | y}^{-1} - \Sigma_{x | y}^{-1}\Sigma_{xy}\left( \Sigma_{yy} + \Sigma_{xy}^{T}\Sigma_{x | y}^{-1}\Sigma_{xy} \right)^{-1}\Sigma_{xy}^{T}\Sigma_{x | y}^{-1} = \Sigma_{x | y}^{-1} - \Sigma_{x | y}^{-1}\alpha\left( \Sigma_{yy}^{-1} + \alpha\Sigma_{x | y}^{-1}\alpha^{T} \right)^{-1}\alpha\Sigma_{x | y}^{-1}\)
\end{enumerate}

\hypertarget{estimators}{%
\section{Estimators}\label{estimators}}

\hypertarget{assume-ridge-beta-optimize-in-sample-prediction-error}{%
\subsection{\texorpdfstring{Assume ridge \(\beta\), optimize in-sample
prediction
error}{Assume ridge \textbackslash{}beta, optimize in-sample prediction error}}\label{assume-ridge-beta-optimize-in-sample-prediction-error}}

\[ \hat{\beta} = \arg\min_{\beta}\left\{\frac{1}{2} \left\| \mathbb{Y} - \mathbb{X}\beta \right\|_{F}^{2} + \frac{\lambda}{2}\left\| \beta \right\|_{F}^{2}\right\} \]

\begin{align*}
  &\Rightarrow -\mathbb{X}^{T}\left( \mathbb{Y} - \mathbb{X}\hat{\beta} \right) + \lambda\hat{\beta} = 0 \\
  & \\
  &\Rightarrow \hat{\beta} = \left( \mathbb{X}^{T}\mathbb{X} + \lambda I_{p} \right)^{-1}\mathbb{X}^{T}\mathbb{Y} = \mathbb{X}^{T}\left( \mathbb{X}\mathbb{X}^{T} + \lambda I_{n} \right)^{-1}\mathbb{Y}
\end{align*}

\hypertarget{assume-ridge-beta-optimize-the-likelihood-for-beta}{%
\subsection{\texorpdfstring{Assume ridge \(\beta\), optimize the
likelihood for
\(\beta\)}{Assume ridge \textbackslash{}beta, optimize the likelihood for \textbackslash{}beta}}\label{assume-ridge-beta-optimize-the-likelihood-for-beta}}

\[ \hat{\beta} = \arg\min_{\beta}\left\{ \frac{1}{n} Tr\left[ \left(\mathbb{Y} - \mathbb{X}\beta \right)^{T}\left( \mathbb{Y} - \mathbb{X}\beta \right)\Sigma_{y | x}^{-1} \right] - \frac{1}{2}log\left| \Sigma_{y | x}^{-1} \right| + \frac{\lambda}{n}\left\| \beta \right\|_{F}^{2} \right\} \]

\begin{align*}
  &\Rightarrow -\frac{2}{n}\mathbb{X}^{T}\left( \mathbb{Y} - \mathbb{X}\hat{\beta} \right)\hat{\Sigma}_{y | x}^{-1} + \frac{2}{n}\lambda\hat{\beta} = 0 \Rightarrow \left( \hat{\Sigma}_{y | x}^{-1} \otimes \mathbb{X}^{T}\mathbb{X} \right)vec\left( \hat{\beta} \right) + \lambda vec\left( \hat{\beta} \right) = vec\left( \mathbb{X}^{T}\mathbb{Y}\hat{\Sigma}_{y | x}^{-1} \right) \\
  & \\
  &\Rightarrow vec\left( \hat{\beta} \right) = \left( \hat{\Sigma}_{y | x}^{-1} \otimes \mathbb{X}^{T}\mathbb{X} + \lambda I_{pr} \right)^{-1}vec\left( \mathbb{X}^{T}\mathbb{Y}\hat{\Sigma}_{y | x}^{-1} \right) \\
\end{align*}

where
\(\hat{\Sigma}_{y | x} = (\mathbb{Y} - \mathbb{X}\hat{\beta} )^{T}(\mathbb{Y} - \mathbb{X}\hat{\beta})/n\).
In order to solve for \(\hat{\beta}\), we need to iterate between
\(\hat{\beta}\) and \(\hat{\Sigma}_{y | x}^{-1}\) until convergence.

Note that if \(r = 1\) then \(\sigma_{y}^{2}\) can be absorbed into
\(\tilde{\lambda}\) and

\[ \hat{\beta} = \left( \mathbb{X}^{T}\mathbb{X} + \tilde{\lambda}I_{p} \right)^{-1}\mathbb{X}^{T}\mathbb{Y} = \mathbb{X}^{T}\left( \mathbb{X}\mathbb{X}^{T} + \tilde{\lambda}_{n}I_{n} \right)^{-1}\mathbb{Y} \]

\hypertarget{assume-ridge-sigma_xx-1-optimize-in-sample-prediction-error}{%
\subsection{\texorpdfstring{Assume ridge \(\Sigma_{xx}^{-1}\), optimize
in-sample prediction
error}{Assume ridge \textbackslash{}Sigma\_\{xx\}\^{}\{-1\}, optimize in-sample prediction error}}\label{assume-ridge-sigma_xx-1-optimize-in-sample-prediction-error}}

\begin{align*}
  \hat{\Sigma}_{xx}^{-1} &= \arg\min_{\Sigma_{xx}^{-1}}\left\{ \frac{1}{2}\left\| \mathbb{Y} - \mathbb{X}\mathbb{\beta} \right\|_{F}^{2} + \frac{\lambda}{2}\left\| \Sigma_{xx}^{-1} \right\|_{F}^{2} \right\} \\
  & \\
  &= \arg\min_{\Sigma_{xx}^{-1}}\left\{ \frac{1}{2}\left\| \mathbb{Y} - \mathbb{X}\Sigma_{xx}^{-1}\Sigma_{xy} \right\|_{F}^{2} + \frac{\lambda}{2}\left\| \Sigma_{xx}^{-1} \right\|_{F}^{2} \right\} \\
\end{align*}

\begin{align*}
  \nabla_{\Sigma_{xx}^{-1}}&\left\{ \frac{1}{2}Tr\left[ \left(\mathbb{Y} - \mathbb{X}\Sigma_{xx}^{-1}\Sigma_{xy} \right)^{T}\left(\mathbb{Y} - \mathbb{X}\Sigma_{xx}^{-1}\Sigma_{xy} \right) \right] + \frac{\lambda}{2}Tr\left[ \Sigma_{xx}^{-1}\Sigma_{xx}^{-1} \right] \right\} \\
  & \\
  &= -\mathbb{X}^{T}\mathbb{Y}\Sigma_{xy}^{T} + \mathbb{X}^{T}\mathbb{X}\Sigma_{xx}^{-1}\Sigma_{xy}\Sigma_{xy}^{T} + \lambda\Sigma_{xx}^{-1} \\
  & \\
  &\Rightarrow vec\left( \hat{\Sigma}_{xx}^{-1} \right) = \left(\hat{\Sigma}_{xy}\hat{\Sigma}_{xy}^{T} \otimes \mathbb{X}^{T}\mathbb{X} + \lambda I_{pp} \right)^{-1}vec\left( \mathbb{X}^{T}\mathbb{Y}\hat{\Sigma}_{xy}^{T} \right) \\
\end{align*}

If we use the sample estimate
\(\hat{\Sigma}_{xy} = \mathbb{X}^{T}\mathbb{Y}/n\) then

\[ vec\left( \hat{\Sigma}_{xx}^{-1} \right) = \left( \mathbb{X}^{T}\mathbb{Y}\mathbb{Y}^{T}\mathbb{X} \otimes \frac{1}{n}\mathbb{X}^{T}\mathbb{X} + \tilde{\lambda}I_{pp} \right)^{-1}vec\left( \mathbb{X}^{T}\mathbb{Y}\mathbb{Y}^{T}\mathbb{X} \right) \]

so that
\(\hat{\beta} = \hat{\Sigma}_{xx}^{-1}\hat{\Sigma}_{xy} = \hat{\Sigma}_{xx}^{-1}\mathbb{X}^{T}\mathbb{Y}/n\).

\hypertarget{assume-ridge-sigma_xx-1-optimize-likelihood-for-beta}{%
\subsection{\texorpdfstring{Assume ridge \(\Sigma_{xx}^{-1}\), optimize
likelihood for
\(\beta\)}{Assume ridge \textbackslash{}Sigma\_\{xx\}\^{}\{-1\}, optimize likelihood for \textbackslash{}beta}}\label{assume-ridge-sigma_xx-1-optimize-likelihood-for-beta}}

\begin{align*}
  \hat{\Sigma}_{xx}^{-1} &= \arg\min_{\Sigma_{xx}^{-1}}\left\{ \frac{1}{n} Tr\left[ \left(\mathbb{Y} - \mathbb{X}\beta \right)^{T}\left( \mathbb{Y} - \mathbb{X}\beta \right)\Sigma_{y | x}^{-1} \right] - \frac{1}{2}log\left| \Sigma_{y | x}^{-1} \right| + \frac{\lambda}{n}\left\| \Sigma_{xx}^{-1} \right\|_{F}^{2} \right\} \\
  & \\
  &= \arg\min_{\Sigma_{xx}^{-1}}\left\{ \frac{1}{n} Tr\left[ \left(\mathbb{Y} - \mathbb{X}\Sigma_{xx}^{-1}\Sigma_{xy} \right)^{T}\left( \mathbb{Y} - \mathbb{X}\Sigma_{xx}^{-1}\Sigma_{xy} \right)\Sigma_{y | x}^{-1} \right] - \frac{1}{2}log\left| \Sigma_{y | x}^{-1} \right| + \frac{\lambda}{n}\left\| \Sigma_{xx}^{-1} \right\|_{F}^{2} \right\} \\
\end{align*}

\begin{align*}
  \nabla_{\Sigma_{xx}^{-1}}&\left\{ \frac{1}{n}Tr\left[ \left(\mathbb{Y} - \mathbb{X}\Sigma_{xx}^{-1}\Sigma_{xy} \right)^{T}\left(\mathbb{Y} - \mathbb{X}\Sigma_{xx}^{-1}\Sigma_{xy} \right)\Sigma_{y | x}^{-1} \right] + \frac{\lambda}{n}Tr\left[ \Sigma_{xx}^{-1}\Sigma_{xx}^{-1} \right] \right\} \\
  & \\
  &= -\frac{2}{n}\mathbb{X}^{T}\mathbb{Y}\Sigma_{y | x}^{-1}\Sigma_{xy}^{T} + \frac{2}{n}\mathbb{X}^{T}\mathbb{X}\Sigma_{xx}^{-1}\Sigma_{xy}\Sigma_{y | x}^{-1}\Sigma_{xy}^{T} + \frac{2\lambda}{n}\Sigma_{xx}^{-1} \\
  & \\
  &\Rightarrow vec\left( \hat{\Sigma}_{xx}^{-1} \right) = \left(\hat{\Sigma}_{xy}\hat{\Sigma}_{y | x}^{-1}\hat{\Sigma}_{xy}^{T} \otimes \mathbb{X}^{T}\mathbb{X} + \lambda I_{pp} \right)^{-1}vec\left( \mathbb{X}^{T}\mathbb{Y}\hat{\Sigma}_{y | x}^{-1}\hat{\Sigma}_{xy}^{T} \right) \\
\end{align*}

If we use the sample estimate
\(\hat{\Sigma}_{xy} = \mathbb{X}^{T}\mathbb{Y}/n\) then

\[ vec\left( \hat{\Sigma}_{xx}^{-1} \right) = \left( \mathbb{X}^{T}\mathbb{Y}\hat{\Sigma}_{y | x}^{-1}\mathbb{Y}^{T}\mathbb{X} \otimes \frac{1}{n}\mathbb{X}^{T}\mathbb{X} + \tilde{\lambda}I_{pp} \right)^{-1}vec\left( \mathbb{X}^{T}\mathbb{Y}\hat{\Sigma}_{y | x}^{-1}\mathbb{Y}^{T}\mathbb{X} \right) \]

where
\(\hat{\Sigma}_{y | x} = (\mathbb{Y} - \mathbb{X}\hat{\beta} )^{T}(\mathbb{Y} - \mathbb{X}\hat{\beta})/n\)
so that in order to solve for
\(\hat{\beta} = \hat{\Sigma}_{xx}^{-1}\hat{\Sigma}_{xy} = \hat{\Sigma}_{xx}^{-1}\mathbb{X}^{T}\mathbb{Y}/n\),
we need to iterate between \(\hat{\beta}\) and
\(\hat{\Sigma}_{y | x}^{-1}\) until convergence.

\hypertarget{assume-ridge-sigma_xx-1-optimize-likelihood-for-sigma_xx-1}{%
\subsection{\texorpdfstring{Assume ridge \(\Sigma_{xx}^{-1}\), optimize
likelihood for
\(\Sigma_{xx}^{-1}\)}{Assume ridge \textbackslash{}Sigma\_\{xx\}\^{}\{-1\}, optimize likelihood for \textbackslash{}Sigma\_\{xx\}\^{}\{-1\}}}\label{assume-ridge-sigma_xx-1-optimize-likelihood-for-sigma_xx-1}}

\[ \hat{\Sigma}_{xx}^{-1} = \arg\min_{\Sigma_{xx}^{-1}}\left\{ Tr\left( S_{xx}\Sigma_{xx}^{-1} \right) - log\left| \Sigma_{xx}^{-1} \right| + \frac{\lambda}{2}\left\| \Sigma_{xx}^{-1} \right\|_{F}^{2} \right\} \]

\begin{align*}
  \nabla_{\Sigma_{xx}^{-1}}&\left\{ Tr\left( S_{xx}\Sigma_{xx}^{-1} \right) - \log\det\left( \Sigma_{xx}^{-1} \right) + \frac{\lambda}{2}\left\| \Sigma_{xx}^{-1} \right\|_{F}^{2} \right\} = S_{xx} - \Sigma_{xx} + \lambda\Sigma_{xx}^{-1} \\
\end{align*}

where \(S_{xx} = \mathbb{X}^{T}\mathbb{X}/n\). Set the gradient equal to
zero and decompose \(\Sigma_{xx}^{-1} = VDV^{T}\) where \(D\) is a
diagonal matrix with diagonal elements equal to the eigen values of
\(\Sigma_{xx}^{-1}\) and \(V\) is the matrix with corresponding eigen
vectors as columns.

\[ S_{xx} = \Sigma_{xx} - \lambda\Sigma_{xx}^{-1} = VD^{-1}V^{T} - \lambda VDV^{T} = V\left(D^{-1} - \lambda D \right)V{^T} \]

\[ \phi_{j}\left( S_{xx} \right) = \frac{1}{\phi_{j}(\Sigma_{xx}^{-1})} - \lambda\phi_{j}\left( \Sigma_{xx}^{-1} \right) \]

\begin{align*}
  &\Rightarrow \lambda\phi_{j}^{2}\left( \Sigma_{xx}^{-1} \right) + \phi_{j}\left( S_{xx} \right)\phi_{j}\left( \Sigma_{xx}^{-1} \right) - 1 = 0 \\
  & \\
  &\Rightarrow \phi_{j}\left( \Sigma_{xx}^{-1} \right) = \frac{-\phi_{j}\left( S_{xx} \right) \pm \sqrt{\phi_{j}^{2}\left( S_{xx} \right) + 4\lambda}}{2\lambda}
\end{align*}

In summary, if we decompose \(S_{xx} = VQV^{T}\) then

\[ \hat{\Sigma}_{xx}^{-1} = \frac{1}{2\lambda}V\left[ -Q + \left( Q^{2} + 4\lambda I_{p} \right)^{1/2} \right]V^{T} \]

so that
\(\hat{\beta} = \hat{\Sigma}_{xx}^{-1}\Sigma_{xy} = \hat{\Sigma}_{xx}^{-1}\mathbb{X}^{T}\mathbb{Y}/n\).

\hypertarget{assume-ridge-sigma_xy-optimize-in-sample-prediction-error}{%
\subsection{\texorpdfstring{Assume ridge \(\Sigma_{xy}\), optimize
in-sample prediction
error}{Assume ridge \textbackslash{}Sigma\_\{xy\}, optimize in-sample prediction error}}\label{assume-ridge-sigma_xy-optimize-in-sample-prediction-error}}

\begin{align*}
  \hat{\Sigma}_{xy} &= \arg\min_{\Sigma_{xy}}\left\{ \frac{1}{2}\left\| \mathbb{Y} - \mathbb{X}\mathbb{\beta} \right\|_{F}^{2} + \frac{\lambda}{2}\left\| \Sigma_{xy} \right\|_{F}^{2} \right\} \\
  & \\
  &= \arg\min_{\Sigma_{xy}}\left\{ \frac{1}{2}\left\| \mathbb{Y} - \mathbb{X}\Sigma_{xx}^{-1}\Sigma_{xy} \right\|_{F}^{2} + \frac{\lambda}{2}\left\| \Sigma_{xy} \right\|_{F}^{2} \right\} \\
\end{align*}

\begin{align*}
  \nabla_{\Sigma_{xy}}&\left\{ \frac{1}{2}Tr\left[ \left(\mathbb{Y} - \mathbb{X}\Sigma_{xx}^{-1}\Sigma_{xy} \right)^{T}\left(\mathbb{Y} - \mathbb{X}\Sigma_{xx}^{-1}\Sigma_{xy} \right) \right] + \frac{\lambda}{2}Tr\left[ \Sigma_{xy}\Sigma_{xy} \right] \right\} \\
  & \\
  &= -\Sigma_{xx}^{-1}\mathbb{X}^{T}\mathbb{Y} + \Sigma_{xx}^{-1}\mathbb{X}^{T}\mathbb{X}\Sigma_{xx}^{-1}\Sigma_{xy} + \lambda\Sigma_{xy} \\
  & \\
  &\Rightarrow \hat{\Sigma}_{xy} = \left(\hat{\Sigma}_{xx}^{-1}\mathbb{X}^{T}\mathbb{X}\hat{\Sigma}_{xx}^{-1} + \lambda I_{p} \right)^{-1}\hat{\Sigma}_{xx}^{-1}\mathbb{X}^{T}\mathbb{Y} \\
\end{align*}

If we use the sample estimate
\(\hat{\Sigma}_{xx} = \mathbb{X}^{T}\mathbb{X}/n\) (assuming
\(\mathbb{X}^{T}\mathbb{X}\) is positive definite) then

\[ \hat{\Sigma}_{xy} = \left[\left(\mathbb{X}^{T}\mathbb{X} \right)^{-1} + \tilde{\lambda}I_{p} \right]^{-1}\left(\mathbb{X}^{T}\mathbb{X} \right)^{-1}\mathbb{X}^{T}\mathbb{Y}/n = \left( I_{p} + \tilde{\lambda}\mathbb{X}^{T}\mathbb{X} \right)^{-1}\mathbb{X}^{T}\mathbb{Y}/n \]

so that

\begin{align*}
  \hat{\beta} &= \left( \mathbb{X}^{T}\mathbb{X} \right)^{-1}\left[ \left( \mathbb{X}^{T}\mathbb{X} \right)^{-1} + \tilde{\lambda}I_{p} \right]^{-1}\left( \mathbb{X}^{T}\mathbb{X} \right)^{-1}\mathbb{X}^{T}\mathbb{Y} \\
  & \\
  &= \left[ \mathbb{X}^{T}\mathbb{X}\left( I_{p} + \tilde{\lambda}\mathbb{X}^{T}\mathbb{X} \right) \right]^{-1}\mathbb{X}^{T}\mathbb{X} = \left( \mathbb{X}^{T}\mathbb{X} + \tilde{\lambda}\mathbb{X}^{T}\mathbb{X}\mathbb{X}^{T}\mathbb{X} \right)^{-1}\mathbb{X}^{T}\mathbb{Y}
\end{align*}

\hypertarget{assume-ridge-sigma_xy-optimize-likelihood-for-beta}{%
\subsection{\texorpdfstring{Assume ridge \(\Sigma_{xy}\), optimize
likelihood for
\(\beta\)}{Assume ridge \textbackslash{}Sigma\_\{xy\}, optimize likelihood for \textbackslash{}beta}}\label{assume-ridge-sigma_xy-optimize-likelihood-for-beta}}

\begin{align*}
  \hat{\Sigma}_{xy} &= \arg\min_{\Sigma_{xy}}\left\{ \frac{1}{n} Tr\left[ \left(\mathbb{Y} - \mathbb{X}\beta \right)^{T}\left( \mathbb{Y} - \mathbb{X}\beta \right)\Sigma_{y | x}^{-1} \right] - \frac{1}{2}log\left| \Sigma_{y | x}^{-1} \right| + \frac{\lambda}{n}\left\| \Sigma_{xy} \right\|_{F}^{2} \right\} \\
  & \\
  &= \arg\min_{\Sigma_{xy}}\left\{ \frac{1}{n} Tr\left[ \left(\mathbb{Y} - \mathbb{X}\Sigma_{xx}^{-1}\Sigma_{xy} \right)^{T}\left( \mathbb{Y} - \mathbb{X}\Sigma_{xx}^{-1}\Sigma_{xy} \right)\Sigma_{y | x}^{-1} \right] - \frac{1}{2}log\left| \Sigma_{y | x}^{-1} \right| + \frac{\lambda}{n}\left\| \Sigma_{xy} \right\|_{F}^{2} \right\} \\
\end{align*}

\begin{align*}
  \nabla_{\Sigma_{xy}}&\left\{ \frac{1}{n}Tr\left[ \left(\mathbb{Y} - \mathbb{X}\Sigma_{xx}^{-1}\Sigma_{xy} \right)^{T}\left(\mathbb{Y} - \mathbb{X}\Sigma_{xx}^{-1}\Sigma_{xy} \right)\Sigma_{y | x}^{-1} \right] + \frac{\lambda}{n}Tr\left[ \Sigma_{xy}^{T}\Sigma_{xy} \right] \right\} \\
  & \\
  &= -\frac{2}{n}\Sigma_{xx}^{-1}\mathbb{X}^{T}\mathbb{Y}\Sigma_{y | x}^{-1} + \frac{2}{n}\Sigma_{xx}^{-1}\mathbb{X}^{T}\mathbb{X}\Sigma_{xx}^{-1}\Sigma_{xy}\Sigma_{y | x}^{-1} + \frac{2\lambda}{n}\Sigma_{xy} \\
\end{align*}

If we use the sample estimate
\(\hat{\Sigma}_{xx} = \mathbb{X}^{T}\mathbb{X}/n\) (assuming
\(\mathbb{X}^{T}\mathbb{X}\) is positive definite) then

\begin{align*}
  vec\left( \hat{\Sigma}_{xy} \right) &= \left[\hat{\Sigma}_{y | x}^{-1} \otimes n\left( \mathbb{X}^{T}\mathbb{X} \right)^{-1} + \tilde{\lambda} I_{rp} \right]^{-1}vec\left[ \left( \mathbb{X}^{T}\mathbb{X} \right)^{-1}\mathbb{X}^{T}\mathbb{Y}\hat{\Sigma}_{y | x}^{-1} \right] \\
  & \\
  &= \left[ \Sigma_{y | x}^{-1} \otimes \left( \tilde{\lambda}\mathbb{X}^{T}\mathbb{X} + I_{p} \right) \right]^{-1}vec\left[ \mathbb{X}^{T}\mathbb{Y}\Sigma_{y | x}^{-1} \right]
\end{align*}

where
\(\hat{\Sigma}_{y | x} = (\mathbb{Y} - \mathbb{X}\hat{\beta} )^{T}(\mathbb{Y} - \mathbb{X}\hat{\beta})/n\)
so that in order to solve for
\(\hat{\beta} = \hat{\Sigma}_{xx}^{-1}\hat{\Sigma}_{xy} = n\left( \mathbb{X}^{T}\mathbb{X} \right)^{-1}\hat{\Sigma}_{xy}\),
we need to iterate between \(\hat{\beta}\) and
\(\hat{\Sigma}_{y | x}^{-1}\) until convergence.

\hypertarget{assume-ridge-sigma_x-y-1-optimize-likelihood-for-sigma_x-y-1}{%
\subsection{\texorpdfstring{Assume ridge \(\Sigma_{x | y}^{-1}\),
optimize likelihood for
\(\Sigma_{x | y}^{-1}\)}{Assume ridge \textbackslash{}Sigma\_\{x \textbar{} y\}\^{}\{-1\}, optimize likelihood for \textbackslash{}Sigma\_\{x \textbar{} y\}\^{}\{-1\}}}\label{assume-ridge-sigma_x-y-1-optimize-likelihood-for-sigma_x-y-1}}

\[ \hat{\Sigma}_{x | y}^{-1} = \arg\min_{\Sigma_{x | y}^{-1}}\left\{ Tr\left( S_{x | y}\Sigma_{x | y}^{-1} \right) - log\left| \Sigma_{x | y}^{-1} \right| + \frac{\lambda}{2}\left\| \Sigma_{x | y}^{-1} \right\|_{F}^{2} \right\} \]

\begin{align*}
  \nabla_{\Sigma_{x | y}^{-1}}&\left\{ Tr\left( S_{x | y}\Sigma_{x | y}^{-1} \right) - \log\det\left( \Sigma_{x | y}^{-1} \right) + \frac{\lambda}{2}\left\| \Sigma_{x | y}^{-1} \right\|_{F}^{2} \right\} = S_{x | y} - \Sigma_{x | y} + \lambda\Sigma_{x | y}^{-1} \\
\end{align*}

where
\(S_{x | y} = \left( \mathbb{X} - \mathbb{Y}\hat{\alpha} \right)^{T}\left( \mathbb{X} - \mathbb{Y}\hat{\alpha} \right)/n\)
and
\(\hat{\alpha} = \left( \mathbb{Y}^{T}\mathbb{Y} \right)^{-1}\mathbb{Y}^{T}\mathbb{X}\).
Following the derivations in section (3.5), if we decompose
\(S_{x | y} = VQV^{T}\) then

\[ \hat{\Sigma}_{x | y}^{-1} = \frac{1}{2\lambda}V\left[ -Q + \left( Q^{2} + 4\lambda I_{p} \right)^{1/2} \right]V^{T} \]

Using the Woodbury Identity, we find that

\[ \hat{\beta} = \hat{\Sigma}_{x | y}^{-1}\Sigma_{xy}\left( \Sigma_{yy} + \Sigma_{xy}^{T}\hat{\Sigma}_{x | y}^{-1}\Sigma_{xy} \right)^{-1}\Sigma_{yy} = \hat{\Sigma}_{x | y}^{-1}\hat{\alpha}^{T}\left( \Sigma_{yy}^{-1} + \hat{\alpha}\hat{\Sigma}_{x | y}^{-1}\hat{\alpha} \right)^{-1} \]

\hypertarget{assume-ridge-alpha-optimize-in-sample-prediction-error-for-alpha}{%
\subsection{\texorpdfstring{Assume ridge \(\alpha\), optimize in-sample
prediction error for
\(\alpha\)}{Assume ridge \textbackslash{}alpha, optimize in-sample prediction error for \textbackslash{}alpha}}\label{assume-ridge-alpha-optimize-in-sample-prediction-error-for-alpha}}

\[ \hat{\alpha} = \arg\min_{\alpha}\left\{\frac{1}{2} \left\| \mathbb{X} - \mathbb{Y}\alpha \right\|_{F}^{2} + \frac{\lambda}{2}\left\| \alpha \right\|_{F}^{2}\right\} \]
